---
Title: "Project - H1B - V1"
Author: "Ritesh Kumar Singh and Group"
Date: "2023-04-14"
---

```{r}
#install.packages("nortest")
# install.packages("forcats")
# install.packages("usmap")
# install.packages("readxl")
# install.packages("viridis")
# install.packages("dplyr")
# install.packages("tibble")
# install.packages("ggplot2")
# install.packages("hrbrthemes")
# install.packages("plotrix")
# install.packages("ggplot2", dependencies = TRUE)
# install.packages("ExcelFunctionsR")
#library(nortest)
# library(forcats)
# library(usmap)
# library(readxl)
# library(viridis)
# library(dplyr)
# library(tibble)
# library(ggplot2)
# library(hrbrthemes)
# library(plotrix)
# library(ExcelFunctionsR)
# options(hrbrthemes.loadfonts = TRUE)
```

```{r}
col <- c("CASE_STATUS","RECEIVED_DATE","DECISION_DATE","VISA_CLASS","JOB_TITLE","SOC_TITLE","EMPLOYER_NAME","EMPLOYER_STATE","AGENT_REPRESENTING_EMPLOYER","WAGE","H_1B_DEPENDENT","WILLFUL_VIOLATOR","EMPLOYER_COUNTRY","NEW_EMPLOYMENT","CONTINUED_EMPLOYMENT","CHANGE_PREVIOUS_EMPLOYMENT","NEW_CONCURRENT_EMPLOYMENT","SOC_CODE","TOTAL_WORKER_POSITIONS")
```

```{r}
d22 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/H1Bdata2022.csv")[ ,col]
```
```{r}
d21 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/H1Bdata2021.csv")[ ,col]
```
```{r}
d20 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/H1Bdata2020.csv")[ ,col]
```
```{r}
d19 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/H1Bdata2019.csv")[ ,col]
```
```{r}    
d18 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/H1Bdata2018.csv")[ ,col]
```
```{r}
d17 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/H1Bdata2017.csv")[ ,col]
```

###### DATA PRE-PROCESSING ######

# Filtering data for H-1B VISA_CLASS
```{r}
d17 <- filter(d17,VISA_CLASS=="H-1B")
d18 <- filter(d18,VISA_CLASS=="H-1B")
d19 <- filter(d19,VISA_CLASS=="H-1B")
d20 <- filter(d20,VISA_CLASS=="H-1B")
d21 <- filter(d21,VISA_CLASS=="H-1B")
d22 <- filter(d22,VISA_CLASS=="H-1B")
```

# filtering out NA values from WAGE column
```{r}
d17 <- d17[!is.na(d17$WAGE),]
d18 <- d18[!is.na(d18$WAGE),]
d19 <- d19[!is.na(d19$WAGE),]
d20 <- d20[!is.na(d20$WAGE),]
d21 <- d21[!is.na(d21$WAGE),]
d22 <- d22[!is.na(d22$WAGE),]
```

# Removing NA values from CASE_STATUS column
```{r}
d17 <- d17[!is.na(d17$CASE_STATUS),]
d18 <- d18[!is.na(d18$CASE_STATUS),]
d19 <- d19[!is.na(d19$CASE_STATUS),]
d20 <- d20[!is.na(d20$CASE_STATUS),]
d21 <- d21[!is.na(d21$CASE_STATUS),]
d22 <- d22[!is.na(d22$CASE_STATUS),]
```

# Converting WAGE columnin into numeric 
```{r}
d17$WAGE <- as.numeric(d17$WAGE)
d18$WAGE <- as.numeric(d18$WAGE)
d19$WAGE <- as.numeric(d19$WAGE)
d20$WAGE <- as.numeric(d20$WAGE)
d21$WAGE <- as.numeric(d21$WAGE)
d22$WAGE <- as.numeric(d22$WAGE)
```

# Creating "SOC" column from SOC_CODE by taking first 2 characters to VLOOKUP Industry
```{r}
d17$SOC <- substr (d17$SOC_CODE, 1,2)
d18$SOC <- substr (d18$SOC_CODE, 1,2)
d19$SOC <- substr (d19$SOC_CODE, 1,2)
d20$SOC <- substr (d20$SOC_CODE, 1,2)
d21$SOC <- substr (d21$SOC_CODE, 1,2)
d22$SOC <- substr (d22$SOC_CODE, 1,2)
```

# Filtering SOC based on all 22 Industries available on Labour website
```{r}
d17 <- d17 %>%
 filter(d17$SOC %in% c("11","13","15","17","19","21","23","25","27","29","31","33","35","37","39","41","43","45","47","49","51","53"))
```

```{r}
d18 <- d18 %>%
 filter(d18$SOC %in% c("11","13","15","17","19","21","23","25","27","29","31","33","35","37","39","41","43","45","47","49","51","53"))
```

```{r}
d19 <- d19 %>%
 filter(d19$SOC %in% c("11","13","15","17","19","21","23","25","27","29","31","33","35","37","39","41","43","45","47","49","51","53"))
```

```{r}
d20 <- d20 %>%
 filter(d20$SOC %in% c("11","13","15","17","19","21","23","25","27","29","31","33","35","37","39","41","43","45","47","49","51","53"))
```

```{r}
d21 <- d21 %>%
 filter(d21$SOC %in% c("11","13","15","17","19","21","23","25","27","29","31","33","35","37","39","41","43","45","47","49","51","53"))
```

```{r}
d22 <- d22 %>%
 filter(d22$SOC %in% c("11","13","15","17","19","21","23","25","27","29","31","33","35","37","39","41","43","45","47","49","51","53"))
```

# Importing SOC list 
```{r}
SOC_lookup <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/SOC_Lookup.csv")
```

# Adding Industry column by linking SOC with respective industries from SOC List file uploaded above
```{r}
lookup_codes <- SOC_lookup$Code_SOC
d17$Industry <- SOC_lookup$Industry [ match ( d17$SOC, lookup_codes)]
d18$Industry <- SOC_lookup$Industry [ match ( d18$SOC, lookup_codes)]
d19$Industry <- SOC_lookup$Industry [ match ( d19$SOC, lookup_codes)]
d20$Industry <- SOC_lookup$Industry [ match ( d20$SOC, lookup_codes)]
d21$Industry <- SOC_lookup$Industry [ match ( d21$SOC, lookup_codes)]
d22$Industry <- SOC_lookup$Industry [ match ( d22$SOC, lookup_codes)]

```

# Adding "YEAR" column in every year dataset
```{r}
d17$YEAR <- paste("2017")
d18$YEAR <- paste("2018")
d19$YEAR <- paste("2019")
d20$YEAR <- paste("2020")
d21$YEAR <- paste("2021")
d22$YEAR <- paste("2022")
```

# Filtering USA states from "EMPLOYER_STATE" Column
```{r}
d17 <- d17 %>%
 filter(d17$EMPLOYER_STATE %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

d18 <- d18 %>%
 filter(d18$EMPLOYER_STATE %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

d19 <- d19 %>%
 filter(d19$EMPLOYER_STATE %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

d20 <- d20 %>%
 filter(d20$EMPLOYER_STATE %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

d21 <- d21 %>%
 filter(d21$EMPLOYER_STATE %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

d22 <- d22 %>%
 filter(d22$EMPLOYER_STATE %in% c("AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA","KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ","NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT","VA","WA","WV","WI","WY"))

```

# Adding processing time column as "TimeTaken"
```{r}
d17$TimeTaken <- difftime(as.Date(d17$DECISION_DATE,"%m/%d/%y"),as.Date(d17$RECEIVED_DATE,"%m/%d/%y"), units = "days")
d18$TimeTaken <- difftime(as.Date(d18$DECISION_DATE,"%m/%d/%y"),as.Date(d18$RECEIVED_DATE,"%m/%d/%y"), units = "days")
d19$TimeTaken <- difftime(as.Date(d19$DECISION_DATE,"%m/%d/%y"),as.Date(d19$RECEIVED_DATE,"%m/%d/%y"), units = "days")
d20$TimeTaken <- difftime(as.Date(d20$DECISION_DATE,"%m/%d/%y"),as.Date(d20$RECEIVED_DATE,"%m/%d/%y"), units = "days")
d21$TimeTaken <- difftime(as.Date(d21$DECISION_DATE,"%m/%d/%y"),as.Date(d21$RECEIVED_DATE,"%m/%d/%y"), units = "days")
d22$TimeTaken <- difftime(as.Date(d22$DECISION_DATE,"%m/%d/%y"),as.Date(d22$RECEIVED_DATE,"%m/%d/%y"), units = "days")
```

# Fixing the TimeTaken column in the dataset AND ADDING A NEW COLUMN AS "TimeTaken1"
## "TimeTaken1" will be now used for computing processing time 
```{r}
d17 <- d17[!is.na(d17$TimeTaken),]
d18 <- d18[!is.na(d18$TimeTaken),]
d19 <- d19[!is.na(d19$TimeTaken),]
d20 <- d20[!is.na(d20$TimeTaken),]
d21 <- d21[!is.na(d21$TimeTaken),]
d22 <- d22[!is.na(d22$TimeTaken),]

d17$TimeTaken1 <- as.numeric(d17$TimeTaken)
d18$TimeTaken1 <- as.numeric(d18$TimeTaken)
d19$TimeTaken1 <- as.numeric(d19$TimeTaken)
d20$TimeTaken1 <- as.numeric(d20$TimeTaken)
d21$TimeTaken1 <- as.numeric(d21$TimeTaken)
d22$TimeTaken1 <- as.numeric(d22$TimeTaken)

d17 <- filter(d17,TimeTaken1>0)
d18 <- filter(d18,TimeTaken1>0)
d19 <- filter(d19,TimeTaken1>0)
d20 <- filter(d20,TimeTaken1>0)
d21 <- filter(d21,TimeTaken1>0)
d22 <- filter(d22,TimeTaken1>0)
```
# Filtering Certified and Denied data for all dataset 
```{r}
d17_certified <- filter(d17,CASE_STATUS=="CERTIFIED")
d18_certified <- filter(d18,CASE_STATUS=="CERTIFIED")
d19_certified <- filter(d19,CASE_STATUS=="CERTIFIED")
d20_certified <- filter(d20,CASE_STATUS=="Certified")
d21_certified <- filter(d21,CASE_STATUS=="Certified")
d22_certified <- filter(d22,CASE_STATUS=="Certified")
d17_denied <- filter(d17,CASE_STATUS=="DENIED")
d18_denied <- filter(d18,CASE_STATUS=="DENIED")
d19_denied <- filter(d19,CASE_STATUS=="DENIED")
d20_denied <- filter(d20,CASE_STATUS=="Denied")
d21_denied <- filter(d21,CASE_STATUS=="Denied")
d22_denied <- filter(d22,CASE_STATUS=="Denied")
```

```{r}
d17_certifiedW <- filter(d17,CASE_STATUS=="CERTIFIED-WITHDRAWN")
d18_certifiedW <- filter(d18,CASE_STATUS=="CERTIFIED-WITHDRAWN")
d19_certifiedW <- filter(d19,CASE_STATUS=="CERTIFIED-WITHDRAWN")
d20_certifiedW <- filter(d20,CASE_STATUS=="Certified - Withdrawn")
d21_certifiedW <- filter(d21,CASE_STATUS=="Certified - Withdrawn")
d22_certifiedW <- filter(d22,CASE_STATUS=="Certified - Withdrawn")
d17_W <- filter(d17,CASE_STATUS=="WITHDRAWN")
d18_W <- filter(d18,CASE_STATUS=="WITHDRAWN")
d19_W <- filter(d19,CASE_STATUS=="WITHDRAWN")
d20_W <- filter(d20,CASE_STATUS=="Withdrawn")
d21_W <- filter(d21,CASE_STATUS=="Withdrawn")
d22_W <- filter(d22,CASE_STATUS=="Withdrawn")
```

# Binding required data
```{r}
AllData <- rbind(d17,d18,d19,d20,d21,d22)
AllData_Cert <- rbind(d17_certified,d18_certified,d19_certified,d20_certified,d21_certified,d22_certified)
AllData_Den <- rbind(d17_denied,d18_denied,d19_denied,d20_denied,d21_denied,d22_denied)
AllData_certifiedW <- rbind(d17_certifiedW,d18_certifiedW,d19_certifiedW,d20_certifiedW,d21_certifiedW,d22_certifiedW)
AllData_W <- rbind(d17_W,d18_W,d19_W,d20_W,d21_W,d22_W)

Precovid <- rbind(d17,d18,d19)
Precovid_Cert <- rbind(d17_certified,d18_certified,d19_certified)
Precovid_Den <- rbind(d17_denied,d18_denied,d19_denied)
Precovid_certifiedW <- rbind(d17_certifiedW,d18_certifiedW,d19_certifiedW)
Precovid_W <- rbind(d17_W,d18_W,d19_W)

Postcovid <- rbind(d20,d21,d22)
Postcovid_Cert <- rbind(d20_certified,d21_certified,d22_certified)
Postcovid_Den <- rbind(d20_denied,d21_denied,d22_denied)
Postcovid_certifiedW <- rbind(d20_certifiedW,d21_certifiedW,d22_certifiedW)
Postcovid_W <- rbind(d20_W,d21_W,d22_W)
```


```{r}
# Assuming the column for industry is called "Industry" and the column for wage is called "Wage"
mean_wage_by_industry <- Precovid_Cert %>% 
  group_by(Industry) %>% 
  summarize(mean_wage = mean(WAGE)) %>% 
  arrange(desc(mean_wage))

# Reorder the factor levels of Industry based on mean wage
mean_wage_by_industry$Industry <- fct_reorder(mean_wage_by_industry$Industry, mean_wage_by_industry$mean_wage)

# Create the bar plot
ggplot(mean_wage_by_industry, aes(x = Industry, y = mean_wage)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  labs(title = "Mean wage by industry", x = "Industry", y = "Mean wage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) # slant the x-axis labels

```
```{r}
# Extract the wage data for Management Occupations
management_wages <- AllData$TimeTaken1[AllData$Industry == "Food Preparation and Serving Related Occupations"]
```

# Normality test for less than 5000 sample data
```{r}
# Perform Shapiro-Wilk test for normality
shapiro.test(management_wages)
```
# Normality test for all type of sample data
```{r}
# Perform Anderson-Darling test for normality
ad.test(management_wages)
```

```{r}
# Perform Anderson-Darling test for each numerical column
for (col in colnames(Postcovid_W)) {
  if (is.numeric(Postcovid_W[[col]])) {
    ad_test <- ad.test(Postcovid_W[[col]])
    if (ad_test$p.value > 0.05) {
      cat(paste(col, "is normally distributed (AD statistic =", round(ad_test$statistic, 4), ", p =", round(ad_test$p.value, 4), ")\n"))
    } else {
      cat(paste(col, "is not normally distributed (AD statistic =", round(ad_test$statistic, 4), ", p =", round(ad_test$p.value, 4), ")\n"))
    }
  }
}
```

```{r}
Precovid_Cert_PSO <- filter(Precovid_Cert,Industry=="Protective Service Occupations")
Precovid_Cert_BGO <- filter(Precovid_Cert,Industry=="Building and Grounds Cleaning and Maintenance Occupations")
```


##### CREATING SAMPLE DATASET #####

```{r}
# # Calculate summary statistics for the Total worker position column
# summary(Precovid_Cert$TOTAL_WORKER_POSITIONS)
# 
# # Plot a histogram for the Total worker position column
# hist(Precovid_Cert$TOTAL_WORKER_POSITIONS)
# 
# # Plot a density function for the Total worker position column
# density(Precovid_Cert$TOTAL_WORKER_POSITIONS)
```

```{r}
# # Generate 10 random samples of size 100 from the "Total worker position" column
# sample_data <- replicate(10, sample(Precovid_Cert$TOTAL_WORKER_POSITIONS, 100))
# # Find the number of observations with a total worker position equal to or greater than 5 in each sample
# count <- apply(sample_data, 2, function(x) sum(x >= 5))
# # Calculate the probability for each sample
# prob <- count / 100
# # Print the probabilities
# print(prob)
```
# Mean and SD Pre covid certified
```{r}
# Assuming that "df" is your data frame and "TimeTaken1" is the column name
mean_time_taken <- mean(Precovid_Den$TimeTaken1, na.rm = TRUE)
sd_time_taken <- sd(Precovid_Den$TimeTaken1, na.rm = TRUE)

# Print the results
cat("Mean of TimeTaken1 column:", mean_time_taken, "\n")
cat("Standard deviation of TimeTaken1 column:", sd_time_taken, "\n")
```
# Mean and SD Post covid certified
```{r}
# Assuming that "df" is your data frame and "TimeTaken1" is the column name
mean_time_taken <- mean(Postcovid_Den$TimeTaken1, na.rm = TRUE)
sd_time_taken <- sd(Postcovid_Den$TimeTaken1, na.rm = TRUE)

# Print the results
cat("Mean of TimeTaken1 column:", mean_time_taken, "\n")
cat("Standard deviation of TimeTaken1 column:", sd_time_taken, "\n")
```
```{r}
table(Precovid_Cert$TimeTaken1, useNA = 'always')
```

## Creating Sample S1 i.e. Pre Covid
```{r}
# # Set the seed for reproducibility
# set.seed(123)
# 
# # Randomly select 5000 rows
# sampled_dataset <- Precovid[sample(nrow(Precovid), 5000),]
# 
# # View the first few rows of the sampled dataset
# head(sampled_dataset)
# 
# # Export the dataset as a CSV file
# write.csv(sampled_dataset, "sampled_dataset_S1.csv", row.names = TRUE)
```
## Creating Sample S2 i.e. Post Covid
```{r}
# # Set the seed for reproducibility
# set.seed(123)
# 
# # Randomly select 5000 rows
# sampled_dataset <- Postcovid[sample(nrow(Postcovid), 5000),]
# 
# # View the first few rows of the sampled dataset
# head(sampled_dataset)
# 
# # Export the dataset as a CSV file
# write.csv(sampled_dataset, "sampled_dataset_S2.csv", row.names = TRUE)
```

# Reading sample data S1 from the file which is Pre covid data
```{r}
sample_data_S1 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/sampled_dataset_S1.csv")
```

# Reading sample data S2 from the file which is Post covid data
```{r}
sample_data_S2 <- read.csv(file = "/Users/ritesh/Library/CloudStorage/OneDrive-SharedLibraries-NortheasternUniversity/Pirasath Rajh Anantharaman Rajagopalan - Project work - drive/Dataset/CSV/sampled_dataset_S2.csv")
```

# Analysing the sample_data_S1 and sample_data_S2
```{r}
# Count the number of different case status pre and post covid
Precovid_count <- count(sample_data_S1, CASE_STATUS)
Postcovid_count <- count(sample_data_S2, CASE_STATUS)

# Calculate the proportion of certified and applications pre and post covid
Precovid_certified_prop <- Precovid_count[1, 2] / sum(Precovid_count$n)
Postcovid_certified_prop <- Postcovid_count[1, 2] / sum(Postcovid_count$n)
```

```{r}
Precovid_certified_prop
Postcovid_certified_prop
```

# Confidence interval for the mean wage in the pre-covid dataset, stratified by case status (certified & denied):
```{r}
pre_covid_certified_mean <- mean(Precovid_Cert$WAGE, na.rm = TRUE)
pre_covid_certified_sd <- sd(Precovid_Cert$WAGE, na.rm = TRUE)

pre_covid_certified_ci <- confint(lm(WAGE ~ 1, data = Precovid_Cert))
pre_covid_certified_lower_ci <- pre_covid_certified_ci[1]
pre_covid_certified_upper_ci <- pre_covid_certified_ci[2]

pre_covid_denied_mean <- mean(Precovid_Den$WAGE)
pre_covid_denied_ci <- confint(lm(WAGE ~ 1, data = Precovid_Den))
pre_covid_denied_lower_ci <- pre_covid_denied_ci[1]
pre_covid_denied_upper_ci <- pre_covid_denied_ci[2]
```

# Confidence interval for the mean wage in the post-covid dataset, stratified by case status (certified & denied):
```{r}
post_covid_certified_mean <- mean(Postcovid_Cert$WAGE, na.rm = TRUE)
post_covid_certified_sd <- sd(Postcovid_Cert$WAGE, na.rm = TRUE)

post_covid_certified_ci <- confint(lm(WAGE ~ 1, data = Postcovid_Cert))
post_covid_certified_lower_ci <- post_covid_certified_ci[1]
post_covid_certified_upper_ci <- post_covid_certified_ci[2]

post_covid_denied_mean <- mean(Postcovid_Den$WAGE)
post_covid_denied_ci <- confint(lm(WAGE ~ 1, data = Postcovid_Den))
post_covid_denied_lower_ci <- post_covid_denied_ci[1]
post_covid_denied_upper_ci <- post_covid_denied_ci[2]
```

## CASE_STATUS values Pre Covid
```{r}
#table(Precovid$CASE_STATUS, useNA = 'always')
table(Precovid$CASE_STATUS)
```
## CASE_STATUS values Post Covid
```{r}
table(Postcovid$CASE_STATUS)
```

##### DIGITAL VISUALIZATION #####

# Classification of Post and Precovid data based on case status
```{r}
specie <- c(rep("Pre covid" , 4) , rep("Post covid" , 4))
Labels <- rep(c("CERTIFIED" , "CERTIFIED-WITHDRAWN" , "DENIED" , "WITHDRAWN") , 2)
value <- c(1659014,113142,19550,56696,1840168,81409,9694,36686)
data <- data.frame(specie,Labels,value)
# ggplot(data, aes(fill=Labels, y=value, x=specie)) + 
#     geom_bar(position="stack", stat="identity") +
#     scale_fill_viridis(discrete = T) +
#     ggtitle("Classification of Post and Precovid data based on case status") +
#     theme_ipsum() +
#     xlab("") + ylab("Cases count")
# Define custom colors
my_colors <- c("#FF6B6B", "#FFC06B", "#8AFF6B", "#6BC4FF")

# Create the plot
ggplot(data, aes(fill=Labels, y=value, x=specie)) + 
    geom_bar(position="stack", stat="identity") +
    scale_fill_manual(values=my_colors) + # Specify custom color palette
    ggtitle("Classification of Post and Precovid data based on case status") +
    theme_ipsum() +
    xlab("") + ylab("Cases count")
```
# To convert exponential value into numerical
```{r}
options(scipen = 999)
```

# Certified case count vs processing time for pre covid data
```{r}
# ggplot(Precovid_Cert, aes(x = TimeTaken1)) +
# geom_histogram(binwidth = 0.5, fill = "steelblue", color = "white") +
# labs(x = "Processing Time", y = "Cases count", title = "Histogram of Processing Time for Pre-covid certified data")

ggplot(Precovid_Cert, aes(x = TimeTaken1)) +
  geom_histogram(binwidth = 0.5, fill = "steelblue", color = "white") +
  labs(x = "Processing Time", y = "Certified case count", title = "Histogram of Processing Time for Pre-covid certified data") +
  scale_x_continuous(breaks = seq(0, max(Precovid_Cert$TimeTaken1), 2))

```
# Certified case count vs processing time for post covid data
```{r}
ggplot(Postcovid_Cert, aes(x = log(TimeTaken1))) +
geom_histogram(binwidth = 0.5, fill = "steelblue", color = "white") +
labs(x = "Natural log of Processing Time", y = "Cases count", title = "Histogram of Natural log of Processing Time for Post covid certified data")
```

```{r}
table(Postcovid_Cert$TimeTaken1, useNA = 'always')
```
